---
title: "R : Exploring Robust Logistic Regression Models for Handling Quasi-Complete Separation"
author: "John Pauline Pineda"
date: "February 19, 2023"
output: 
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
    highlight: tango
    css: doc.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=15, fig.height=10)
```
# **1. Table of Contents**
|
| This document presents a non-exhaustive list of robust model variants applied to handle quasi-complete or complete separation during logistic regression modelling using various helpful packages in <mark style="background-color: #CCECFF">**R**</mark>.    
|
##  1.1 Sample Data
|
| The <mark style="background-color: #EEEEEE;color: #FF0000">**sex2**</mark>  dataset from the  <mark style="background-color: #CCECFF">**logistf**</mark> package was used for this illustrated example. One of the original categorical predictors was transformed to better simulate a quasi-complete condition when a covariate almost perfectly predicts the outcome.
|
| Preliminary dataset assessment:
|
| **[A]** 239 rows (observations)
|      **[A.1]** Train Set = 239 observations
| 
| **[B]** 7 columns (variables)
|      **[B.1]** 1/7 response = <span style="color: #FF0000">UTI</span> variable (factor)
|             **[B.1.1]** Levels = <span style="color: #FF0000">UTI=No</span> < <span style="color: #FF0000">UTI=Yes</span>
|      **[B.2]** 6/7 predictors = All remaining variables (6/6 factor)
|     
| 
```{r section_1.1, warning=FALSE, message=FALSE}
##################################
# Loading R libraries
##################################
library(AppliedPredictiveModeling)
library(caret)
library(rpart)
library(lattice)
library(dplyr)
library(tidyr)
library(moments)
library(skimr)
library(RANN)
library(pls)
library(corrplot)
library(tidyverse)
library(lares)
library(DMwR)
library(gridExtra)
library(rattle)
library(rpart.plot)
library(RColorBrewer)
library(stats)
library(nnet)
library(elasticnet)
library(earth)
library(party)
library(kernlab)
library(randomForest)
library(Cubist)
library(pROC)
library(mda)
library(klaR)
library(pamr)
library(OptimalCutpoints)
library(broom)
library(PRROC)
library(logistf)
library(arm)
library(glmnet)

##################################
# Loading source and
# formulating the train set
##################################
data(sex2)
sex2$dia <- sex2$case
sex2$dia[1] <- 0
UTI_Train <- as.data.frame(sex2)

##################################
# Applying verbose descriptions
# for the response and predictor variables
##################################
colnames(UTI_Train) <- c("UTI",
                         "Age_24",
                         "Contraceptive",
                         "Condom",
                         "LubricatedCondom",
                         "Spermicide",
                         "Diaphragm")

UTI <- UTI_Train

UTI_Train <- lapply(UTI_Train, function(x) ifelse(x==1,"Yes","No"))
UTI_Train <- lapply(UTI_Train, function(x) as.factor(as.character(x)))
UTI_Train <- as.data.frame(UTI_Train)

##################################
# Performing a general exploration of the train set
##################################
dim(UTI_Train)
str(UTI_Train)
summary(UTI_Train)

##################################
# Formulating a data type assessment summary
##################################
PDA <- UTI_Train
(PDA.Summary <- data.frame(
  Column.Index=c(1:length(names(PDA))),
  Column.Name= names(PDA),
  Column.Type=sapply(PDA, function(x) class(x)),
  row.names=NULL)
)
```

##  1.2 Data Quality Assessment
|
| Data quality assessment:
|
| **[A]** No missing observations noted for any variable.
|
| **[B]** Low variance observed for 2 variables with First.Second.Mode.Ratio>5.
|      **[B.1]** <span style="color: #FF0000">Age_24</span> variable (factor)
|
| **[C]** Due to the absence of numeric predictors, no low variance with Unique.Count.Ratio<0.01 and high skewness with Skewness>3 or Skewness<(-3) observed for any variable.
|
|
```{r section_1.2, warning=FALSE, message=FALSE}
##################################
# Loading dataset
##################################
DQA <- UTI_Train

##################################
# Formulating an overall data quality assessment summary
##################################
(DQA.Summary <- data.frame(
  Column.Index=c(1:length(names(DQA))),
  Column.Name= names(DQA),
  Column.Type=sapply(DQA, function(x) class(x)),
  Row.Count=sapply(DQA, function(x) nrow(DQA)),
  NA.Count=sapply(DQA,function(x)sum(is.na(x))),
  Fill.Rate=sapply(DQA,function(x)format(round((sum(!is.na(x))/nrow(DQA)),3),nsmall=3)),
  row.names=NULL)
)

##################################
# Listing all predictors
##################################
DQA.Predictors <- DQA[,!names(DQA) %in% c("UTI")]

##################################
# Listing all numeric predictors
##################################
DQA.Predictors.Numeric <- DQA.Predictors[,sapply(DQA.Predictors, is.numeric)]

if (length(names(DQA.Predictors.Numeric))>0) {
    print(paste0("There are ",
               (length(names(DQA.Predictors.Numeric))),
               " numeric predictor variable(s)."))
} else {
  print("There are no numeric predictor variables.")
}

##################################
# Listing all factor predictors
##################################
DQA.Predictors.Factor <- DQA.Predictors[,sapply(DQA.Predictors, is.factor)]

if (length(names(DQA.Predictors.Factor))>0) {
    print(paste0("There are ",
               (length(names(DQA.Predictors.Factor))),
               " factor predictor variable(s)."))
} else {
  print("There are no factor predictor variables.")
}

##################################
# Formulating a data quality assessment summary for factor predictors
##################################
if (length(names(DQA.Predictors.Factor))>0) {

  ##################################
  # Formulating a function to determine the first mode
  ##################################
  FirstModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    ux[tab == max(tab)]
  }

  ##################################
  # Formulating a function to determine the second mode
  ##################################
  SecondModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    fm = ux[tab == max(tab)]
    sm = x[!(x %in% fm)]
    usm <- unique(sm)
    tabsm <- tabulate(match(sm, usm))
    ifelse(is.na(usm[tabsm == max(tabsm)])==TRUE,
           return("x"),
           return(usm[tabsm == max(tabsm)]))
  }

  (DQA.Predictors.Factor.Summary <- data.frame(
  Column.Name= names(DQA.Predictors.Factor),
  Column.Type=sapply(DQA.Predictors.Factor, function(x) class(x)),
  Unique.Count=sapply(DQA.Predictors.Factor, function(x) length(unique(x))),
  First.Mode.Value=sapply(DQA.Predictors.Factor, function(x) as.character(FirstModes(x)[1])),
  Second.Mode.Value=sapply(DQA.Predictors.Factor, function(x) as.character(SecondModes(x)[1])),
  First.Mode.Count=sapply(DQA.Predictors.Factor, function(x) sum(na.omit(x) == FirstModes(x)[1])),
  Second.Mode.Count=sapply(DQA.Predictors.Factor, function(x) sum(na.omit(x) == SecondModes(x)[1])),
  Unique.Count.Ratio=sapply(DQA.Predictors.Factor, function(x) format(round((length(unique(x))/nrow(DQA.Predictors.Factor)),3), nsmall=3)),
  First.Second.Mode.Ratio=sapply(DQA.Predictors.Factor, function(x) format(round((sum(na.omit(x) == FirstModes(x)[1])/sum(na.omit(x) == SecondModes(x)[1])),3), nsmall=3)),
  row.names=NULL)
  )

}

##################################
# Formulating a data quality assessment summary for numeric predictors
##################################
if (length(names(DQA.Predictors.Numeric))>0) {

  ##################################
  # Formulating a function to determine the first mode
  ##################################
  FirstModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    ux[tab == max(tab)]
  }

  ##################################
  # Formulating a function to determine the second mode
  ##################################
  SecondModes <- function(x) {
    ux <- unique(na.omit(x))
    tab <- tabulate(match(x, ux))
    fm = ux[tab == max(tab)]
    sm = na.omit(x)[!(na.omit(x) %in% fm)]
    usm <- unique(sm)
    tabsm <- tabulate(match(sm, usm))
    ifelse(is.na(usm[tabsm == max(tabsm)])==TRUE,
           return(0.00001),
           return(usm[tabsm == max(tabsm)]))
  }

  (DQA.Predictors.Numeric.Summary <- data.frame(
  Column.Name= names(DQA.Predictors.Numeric),
  Column.Type=sapply(DQA.Predictors.Numeric, function(x) class(x)),
  Unique.Count=sapply(DQA.Predictors.Numeric, function(x) length(unique(x))),
  Unique.Count.Ratio=sapply(DQA.Predictors.Numeric, function(x) format(round((length(unique(x))/nrow(DQA.Predictors.Numeric)),3), nsmall=3)),
  First.Mode.Value=sapply(DQA.Predictors.Numeric, function(x) format(round((FirstModes(x)[1]),3),nsmall=3)),
  Second.Mode.Value=sapply(DQA.Predictors.Numeric, function(x) format(round((SecondModes(x)[1]),3),nsmall=3)),
  First.Mode.Count=sapply(DQA.Predictors.Numeric, function(x) sum(na.omit(x) == FirstModes(x)[1])),
  Second.Mode.Count=sapply(DQA.Predictors.Numeric, function(x) sum(na.omit(x) == SecondModes(x)[1])),
  First.Second.Mode.Ratio=sapply(DQA.Predictors.Numeric, function(x) format(round((sum(na.omit(x) == FirstModes(x)[1])/sum(na.omit(x) == SecondModes(x)[1])),3), nsmall=3)),
  Minimum=sapply(DQA.Predictors.Numeric, function(x) format(round(min(x,na.rm = TRUE),3), nsmall=3)),
  Mean=sapply(DQA.Predictors.Numeric, function(x) format(round(mean(x,na.rm = TRUE),3), nsmall=3)),
  Median=sapply(DQA.Predictors.Numeric, function(x) format(round(median(x,na.rm = TRUE),3), nsmall=3)),
  Maximum=sapply(DQA.Predictors.Numeric, function(x) format(round(max(x,na.rm = TRUE),3), nsmall=3)),
  Skewness=sapply(DQA.Predictors.Numeric, function(x) format(round(skewness(x,na.rm = TRUE),3), nsmall=3)),
  Kurtosis=sapply(DQA.Predictors.Numeric, function(x) format(round(kurtosis(x,na.rm = TRUE),3), nsmall=3)),
  Percentile25th=sapply(DQA.Predictors.Numeric, function(x) format(round(quantile(x,probs=0.25,na.rm = TRUE),3), nsmall=3)),
  Percentile75th=sapply(DQA.Predictors.Numeric, function(x) format(round(quantile(x,probs=0.75,na.rm = TRUE),3), nsmall=3)),
  row.names=NULL)
  )

}

##################################
# Identifying potential data quality issues
##################################

##################################
# Checking for missing observations
##################################
if ((nrow(DQA.Summary[DQA.Summary$NA.Count>0,]))>0){
  print(paste0("Missing observations noted for ",
               (nrow(DQA.Summary[DQA.Summary$NA.Count>0,])),
               " variable(s) with NA.Count>0 and Fill.Rate<1.0."))
  DQA.Summary[DQA.Summary$NA.Count>0,]
} else {
  print("No missing observations noted.")
}

##################################
# Checking for zero or near-zero variance predictors
##################################
if (length(names(DQA.Predictors.Factor))==0) {
  print("No factor predictors noted.")
} else if (nrow(DQA.Predictors.Factor.Summary[as.numeric(as.character(DQA.Predictors.Factor.Summary$First.Second.Mode.Ratio))>5,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Predictors.Factor.Summary[as.numeric(as.character(DQA.Predictors.Factor.Summary$First.Second.Mode.Ratio))>5,])),
               " factor variable(s) with First.Second.Mode.Ratio>5."))
  DQA.Predictors.Factor.Summary[as.numeric(as.character(DQA.Predictors.Factor.Summary$First.Second.Mode.Ratio))>5,]
} else {
  print("No low variance factor predictors due to high first-second mode ratio noted.")
}

if (length(names(DQA.Predictors.Numeric))==0) {
  print("No numeric predictors noted.")
} else if (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$First.Second.Mode.Ratio))>5,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$First.Second.Mode.Ratio))>5,])),
               " numeric variable(s) with First.Second.Mode.Ratio>5."))
  DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$First.Second.Mode.Ratio))>5,]
} else {
  print("No low variance numeric predictors due to high first-second mode ratio noted.")
}

if (length(names(DQA.Predictors.Numeric))==0) {
  print("No numeric predictors noted.")
} else if (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Unique.Count.Ratio))<0.01,])>0){
  print(paste0("Low variance observed for ",
               (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Unique.Count.Ratio))<0.01,])),
               " numeric variable(s) with Unique.Count.Ratio<0.01."))
  DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Unique.Count.Ratio))<0.01,]
} else {
  print("No low variance numeric predictors due to low unique count ratio noted.")
}

##################################
# Checking for skewed predictors
##################################
if (length(names(DQA.Predictors.Numeric))==0) {
  print("No numeric predictors noted.")
} else if (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))>3 |
                                               as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))<(-3),])>0){
  print(paste0("High skewness observed for ",
  (nrow(DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))>3 |
                                               as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))<(-3),])),
  " numeric variable(s) with Skewness>3 or Skewness<(-3)."))
  DQA.Predictors.Numeric.Summary[as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))>3 |
                                 as.numeric(as.character(DQA.Predictors.Numeric.Summary$Skewness))<(-3),]
} else {
  print("No skewed numeric predictors noted.")
}

```

##  1.3 Data Preprocessing

###  1.3.1 Pre-Processed Dataset
|
| Preliminary dataset assessment:
|
| **[A]** 239 rows (observations)
|      **[A.1]** Train Set = 239 observations
| 
| **[B]** 7 columns (variables)
|      **[B.1]** 1/7 response = <span style="color: #FF0000">UTI</span> variable (factor)
|             **[B.1.1]** Levels = <span style="color: #FF0000">UTI=No</span> < <span style="color: #FF0000">UTI=Yes</span>
|      **[B.2]** 6/7 predictors = All remaining variables (6/6 factor)
| 
| **[C]** Pre-processing actions applied:
|      **[C.1]** Due to the objective of simulating a a quasi-complete separation, the predictors observed with near-zero variance were not removed in the analysis 
|
```{r section_1.3.1, warning=FALSE, message=FALSE}
##################################
# Creating the pre-modelling
# train set
##################################
PMA_PreModelling_Train <- UTI_Train

##################################
# Gathering descriptive statistics
##################################
(PMA_PreModelling_Train_Skimmed <- skim(PMA_PreModelling_Train))

###################################
# Verifying the data dimensions
# for the train set
###################################
dim(PMA_PreModelling_Train)

```

## 1.4 Data Exploration
|
| Exploratory data analysis:
|
| **[A]** Several factor variables demonstrated differential relationships with the <span style="color: #FF0000">UTI</span> response variable:
|      **[A.1]** <span style="color: #FF0000">Diaphragm</span> variable (factor)
|      **[A.2]** <span style="color: #FF0000">Age_24</span> variable (factor)
|      **[A.3]** <span style="color: #FF0000">Spermicide</span> variable (factor)
|      **[A.4]** <span style="color: #FF0000">LubricatedCondom</span> variable (factor)
|
```{r section_1.4, warning=FALSE, message=FALSE}
##################################
# Restructuring the dataset for
# for barchart analysis
##################################
EDA.Bar.Source <- PMA_PreModelling_Train

##################################
# Creating a function to formulate
# the proportions table
##################################
EDA.PropTable.Function <- function(FactorVar) {
  EDA.Bar.Source.FactorVar <- EDA.Bar.Source[,c("UTI",
                                                FactorVar)]
  EDA.Bar.Source.FactorVar.Prop <- as.data.frame(prop.table(table(EDA.Bar.Source.FactorVar), 2))
  names(EDA.Bar.Source.FactorVar.Prop)[2] <- "UTI"
  EDA.Bar.Source.FactorVar.Prop$Variable <- rep(FactorVar,nrow(EDA.Bar.Source.FactorVar.Prop))

  return(EDA.Bar.Source.FactorVar.Prop)

}

EDA.Bar.Source.FactorVar.Prop <- rbind(EDA.PropTable.Function("Age_24"),
                                       EDA.PropTable.Function("Contraceptive"),
                                       EDA.PropTable.Function("Condom"),
                                       EDA.PropTable.Function("LubricatedCondom"),
                                       EDA.PropTable.Function("Spermicide"),
                                       EDA.PropTable.Function("Diaphragm"))

(EDA.Barchart.FactorVar <- barchart(EDA.Bar.Source.FactorVar.Prop[,3] ~
                                      EDA.Bar.Source.FactorVar.Prop[,2] | EDA.Bar.Source.FactorVar.Prop[,4],
                                      data=EDA.Bar.Source.FactorVar.Prop,
                                      groups = EDA.Bar.Source.FactorVar.Prop[,1],
                                      stack=TRUE,
                                      ylab = "Proportion",
                                      xlab = "UTI",
                                      auto.key = list(adj = 1),
                                      layout=(c(3,2))))

```

## 1.5 Predictive Model Index Development and Probability Curve Estimation

###  1.5.1 Logistic Regression (LR)
|
| **[A]** The logistic regression model from the <mark style="background-color: #CCECFF">**stats**</mark> package was implemented. The <span style="color: #FF0000">UTI</span> response was regressed against the <span style="color: #FF0000">Age_24</span>, <span style="color: #FF0000">Contraceptive</span>, <span style="color: #FF0000">Condom</span>, <span style="color: #FF0000">LubricatedCondom</span>, <span style="color: #FF0000">Spermicide</span> and <span style="color: #FF0000">Diaphragm</span> predictors.
|
| **[B]** The model does not contain any hyperparameter.
|
| **[C]** The model was not able to sufficiently compensate for the quasi-complete condition simulated for the <span style="color: #FF0000">Diaphragm</span> predictor, resulting in the hyper-inflation of its estimated coefficients and standard errors.
|      **[C.1]** <span style="color: #FF0000">Intercept</span> coefficient = -2.868 
|      **[C.2]** <span style="color: #FF0000">Age_24</span> coefficient = -1.052
|      **[C.3]** <span style="color: #FF0000">Contraceptive</span> coefficient = -21.885
|      **[C.4]** <span style="color: #FF0000">Condom</span> coefficient = -22.120 
|      **[C.5]** <span style="color: #FF0000">LubricatedCondom</span> coefficient = -1.687
|      **[C.6]** <span style="color: #FF0000">Spermicide</span> coefficient = +2.868
|      **[C.7]** <span style="color: #FF0000">Diaphragm</span> coefficient = +71.283
|
| **[D]** The logistic curve formulated by plotting the predicted probabilities against the classification index using the logit values showed a non-logistic profile with predicted points clustered in groups.
|
| **[E]** The performance of the model is summarized as follows:
|      **[E.1]** Final model configuration is fixed due to the absence of a hyperparameter
|      **[E.2]** Apparent ROC Curve AUC = 0.99996
|      **[E.3]** Estimated probabilities were practically dichotomized into a limited set of values. 
|
```{r section_1.5.1, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_LR <- UTI

##################################
# Formulating the Logistic Regression model
##################################
LR_Model <- glm(UTI ~ Age_24 + Contraceptive + Condom + LubricatedCondom + Spermicide + Diaphragm,
                data = PMA_PreModelling_Train_LR,
                family = binomial)

##################################
# Consolidating the model results
##################################
summary(LR_Model)

##################################
# Computing the model predictions
##################################
(LR_Model_Probabilities <- predict(LR_Model, 
                              type = c("response")))

##################################
# Creating a classification index
# based from the model predictions
##################################
(LR_Model_Indices <- predict(LR_Model, 
                           type = c("link")))
max(LR_Model_Indices)
min(LR_Model_Indices)

##################################
# Consolidating the model probabilities
# and classification index
# based from the model predictions
##################################
LR_Model_Predictions <- as.data.frame(PMA_PreModelling_Train_LR)
LR_Model_Predictions$LR_Prob <- LR_Model_Probabilities
LR_Model_Predictions$LR_LP <- LR_Model_Indices
LR_Model_Predictions$UTI <- as.factor(LR_Model_Predictions$UTI)

##################################
# Formulating the probability curve
# using the consolidated model predictions
##################################
LR_Model_Predictions %>%
  ggplot(aes(x = LR_LP ,
             y = LR_Prob,
             color = UTI)) +
  scale_colour_manual(values=c("#1846BA55","#B8000055")) +
  geom_point(size=5) +
  geom_line(color="black") +
  xlab("UTI Classification Index (Logit Values)") +
  ylab("Estimated UTI Probability") +
  labs(color = "UTI") +
  scale_x_continuous( limits=c(-50,50), breaks=seq(-50,50,by=5)) +
  scale_y_continuous( limits=c(0,1), breaks=seq(0,1,by=0.1),labels = scales::percent) +
  ggtitle("Estimated UTI Probabilities Based on Classification Index : Logistic Regression") +
  theme_light() +
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.50),
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        legend.position="top")

##################################
# Formulating the corresponding
# receiver operating characteristic (ROC) curve
# using the model predictions
##################################
LR_Prob_Low  <- LR_Model_Predictions[LR_Model_Predictions$UTI==0,
                                    c("LR_Prob")]
LR_Prob_High <- LR_Model_Predictions[LR_Model_Predictions$UTI==1,
                                    c("LR_Prob")]
LR_Model_Prob_ROC <- roc.curve(scores.class1 = LR_Prob_Low,
                               scores.class0 = LR_Prob_High,
                                curve = TRUE)

plot(LR_Model_Prob_ROC,
     xlab="1-Specificity", 
     ylab="Sensitivity",
     main="ROC Curve of the UTI Probabilities : Logistic Regression", 
     color=TRUE, 
     lwd=8,
     legend=3)

```

###  1.5.2 Firth's Bias-Reduced Logistic Regression (FBRLR)
|
| **[A]** The Firth's bias-reduced logistic regression model from the <mark style="background-color: #CCECFF">**logistf**</mark> package was implemented. The <span style="color: #FF0000">UTI</span> response was regressed against the <span style="color: #FF0000">Age_24</span>, <span style="color: #FF0000">Contraceptive</span>, <span style="color: #FF0000">Condom</span>, <span style="color: #FF0000">LubricatedCondom</span>, <span style="color: #FF0000">Spermicide</span> and <span style="color: #FF0000">Diaphragm</span> predictors.
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">maxit</span> = maximum number of iterations for the penalized-likelihood logistic regression held constant at a value of 1000
|
| **[C]** The model was not able to sufficiently compensate for the quasi-complete condition simulated for the <span style="color: #FF0000">Diaphragm</span> predictor, resulting in reasonably valid estimated coefficients and standard errors.
|      **[C.1]** <span style="color: #FF0000">Intercept</span> coefficient = -3.345 
|      **[C.2]** <span style="color: #FF0000">Age_24</span> coefficient = -1.539
|      **[C.3]** <span style="color: #FF0000">Contraceptive</span> coefficient = -1.014
|      **[C.4]** <span style="color: #FF0000">Condom</span> coefficient = -2.048 
|      **[C.5]** <span style="color: #FF0000">LubricatedCondom</span> coefficient = -1.214
|      **[C.6]** <span style="color: #FF0000">Spermicide</span> coefficient = +3.019
|      **[C.7]** <span style="color: #FF0000">Diaphragm</span> coefficient = +10.409
|
| **[D]** The logistic curve formulated by plotting the predicted probabilities against the classification index using the logit values showed a valid logistic profile with reasonably distributed predicted points.
|
| **[E]** The performance of the model is summarized as follows:
|      **[E.1]** Final model configuration involves maxit=1000
|      **[E.2]** Apparent ROC Curve AUC = 0.99996
|      **[E.3]** Estimated probabilities were reasonably distributed across a set of values. 
|
```{r section_1.5.2, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_FBRLR <- UTI

##################################
# Formulating the Logistic Regression model
##################################
FBRLR_Model <- logistf(UTI ~ Age_24 + Contraceptive + Condom + LubricatedCondom + Spermicide + Diaphragm,
                       data = PMA_PreModelling_Train_FBRLR,
                       control=logistf.control(maxit = 10000))

##################################
# Consolidating the model results
##################################
summary(FBRLR_Model)

##################################
# Computing the model predictions
##################################
(FBRLR_Model_Probabilities <- predict(FBRLR_Model, 
                              type = c("response")))

##################################
# Creating a classification index
# based from the model predictions
##################################
(FBRLR_Model_Indices <- predict(FBRLR_Model, 
                           type = c("link")))
max(FBRLR_Model_Indices)
min(FBRLR_Model_Indices)

##################################
# Consolidating the model probabilities
# and classification index
# based from the model predictions
##################################
FBRLR_Model_Predictions <- as.data.frame(PMA_PreModelling_Train_FBRLR)
FBRLR_Model_Predictions$FBRLR_Prob <- FBRLR_Model_Probabilities
FBRLR_Model_Predictions$FBRLR_LP <- FBRLR_Model_Indices
FBRLR_Model_Predictions$UTI <- as.factor(FBRLR_Model_Predictions$UTI)

##################################
# Formulating the probability curve
# using the consolidated model predictions
##################################
FBRLR_Model_Predictions %>%
  ggplot(aes(x = FBRLR_LP ,
             y = FBRLR_Prob,
             color = UTI)) +
  scale_colour_manual(values=c("#1846BA55","#B8000055")) +
  geom_point(size=5) +
  geom_line(color="black") +
  xlab("UTI Classification Index (Logit Values)") +
  ylab("Estimated UTI Probability") +
  labs(color = "UTI") +
  scale_x_continuous( limits=c(-10,10), breaks=seq(-10,10,by=1)) +
  scale_y_continuous( limits=c(0,1), breaks=seq(0,1,by=0.1),labels = scales::percent) +
  ggtitle("Estimated UTI Probabilities Based on Classification Index : Firth's Bias-Reduced Logistic Regression") +
  theme_light() +
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.50),
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        legend.position="top")

##################################
# Formulating the corresponding
# receiver operating characteristic (ROC) curve
# using the model predictions
##################################
FBRLR_Prob_Low  <- FBRLR_Model_Predictions[FBRLR_Model_Predictions$UTI==0,
                                    c("FBRLR_Prob")]
FBRLR_Prob_High <- FBRLR_Model_Predictions[FBRLR_Model_Predictions$UTI==1,
                                    c("FBRLR_Prob")]
FBRLR_Model_Prob_ROC <- roc.curve(scores.class1 = FBRLR_Prob_Low,
                               scores.class0 = FBRLR_Prob_High,
                                curve = TRUE)

plot(FBRLR_Model_Prob_ROC,
     xlab="1-Specificity", 
     ylab="Sensitivity",
     main="ROC Curve of the UTI Probabilities : Firth's Bias-Reduced Logistic Regression", 
     color=TRUE, 
     lwd=8,
     legend=3)

```

###  1.5.3 Firth's Logistic Regression With Added Covariate (FLAC)
|
| **[A]** The Firth's logistic regression with added covariate model from the <mark style="background-color: #CCECFF">**logistf**</mark> package was implemented. The <span style="color: #FF0000">UTI</span> response was regressed against the <span style="color: #FF0000">Age_24</span>, <span style="color: #FF0000">Contraceptive</span>, <span style="color: #FF0000">Condom</span>, <span style="color: #FF0000">LubricatedCondom</span>, <span style="color: #FF0000">Spermicide</span> and <span style="color: #FF0000">Diaphragm</span> predictors.
|
| **[B]** The model contains 1 hyperparameter:
|      **[B.1]** <span style="color: #FF0000">maxit</span> = maximum number of iterations for the penalized-likelihood logistic regression held constant at a value of 1000
|
| **[C]** The model was not able to sufficiently compensate for the quasi-complete condition simulated for the <span style="color: #FF0000">Diaphragm</span> predictor, resulting in reasonably valid estimated coefficients and standard errors.
|      **[C.1]** <span style="color: #FF0000">Intercept</span> coefficient = -3.302 
|      **[C.2]** <span style="color: #FF0000">Age_24</span> coefficient = -1.487
|      **[C.3]** <span style="color: #FF0000">Contraceptive</span> coefficient = -1.042
|      **[C.4]** <span style="color: #FF0000">Condom</span> coefficient = -2.115 
|      **[C.5]** <span style="color: #FF0000">LubricatedCondom</span> coefficient = -1.231
|      **[C.6]** <span style="color: #FF0000">Spermicide</span> coefficient = +3.115
|      **[C.7]** <span style="color: #FF0000">Diaphragm</span> coefficient = +10.507
|
| **[D]** The logistic curve formulated by plotting the predicted probabilities against the classification index using the logit values showed a valid logistic profile with reasonably distributed predicted points.
|
| **[E]** The performance of the model is summarized as follows:
|      **[E.1]** Final model configuration involves maxit=1000
|      **[E.2]** Apparent ROC Curve AUC = 0.99996
|      **[E.3]** Estimated probabilities were reasonably distributed across a set of values. 
|
```{r section_1.5.3, warning=FALSE, message=FALSE}
##################################
# Creating a local object
# for the train set
##################################
PMA_PreModelling_Train_FLAC <- UTI

##################################
# Formulating the Logistic Regression model
##################################
FLAC_Model <- flac(UTI ~ Age_24 + Contraceptive + Condom + LubricatedCondom + Spermicide + Diaphragm,
                       data = PMA_PreModelling_Train_FLAC,
                       control=logistf.control(maxit = 10000))

##################################
# Consolidating the model results
##################################
summary(FLAC_Model)

##################################
# Computing the model predictions
##################################
(FLAC_Model_Probabilities <- predict(FLAC_Model, 
                              type = c("response")))

##################################
# Creating a classification index
# based from the model predictions
##################################
(FLAC_Model_Indices <- predict(FLAC_Model, 
                           type = c("link")))
max(FLAC_Model_Indices)
min(FLAC_Model_Indices)

##################################
# Consolidating the model probabilities
# and classification index
# based from the model predictions
##################################
FLAC_Model_Predictions <- as.data.frame(PMA_PreModelling_Train_FLAC)
FLAC_Model_Predictions$FLAC_Prob <- FLAC_Model_Probabilities
FLAC_Model_Predictions$FLAC_LP <- FLAC_Model_Indices
FLAC_Model_Predictions$UTI <- as.factor(FLAC_Model_Predictions$UTI)

##################################
# Formulating the probability curve
# using the consolidated model predictions
##################################
FLAC_Model_Predictions %>%
  ggplot(aes(x = FLAC_LP ,
             y = FLAC_Prob,
             color = UTI)) +
  scale_colour_manual(values=c("#1846BA55","#B8000055")) +
  geom_point(size=5) +
  geom_line(color="black") +
  xlab("UTI Classification Index (Logit Values)") +
  ylab("Estimated UTI Probability") +
  labs(color = "UTI") +
  scale_x_continuous( limits=c(-10,10), breaks=seq(-10,10,by=1)) +
  scale_y_continuous( limits=c(0,1), breaks=seq(0,1,by=0.1),labels = scales::percent) +
  ggtitle("Estimated UTI Probabilities Based on Classification Index : Firth's Logistic Regression With Added Covariate") +
  theme_light() +
  theme(plot.title = element_text(color="black", size=14, face="bold", hjust=0.50),
        axis.title.x = element_text(color="black", size=12, face="bold"),
        axis.title.y = element_text(color="black", size=12, face="bold"),
        legend.position="top")

##################################
# Formulating the corresponding
# receiver operating characteristic (ROC) curve
# using the model predictions
##################################
FLAC_Prob_Low  <- FLAC_Model_Predictions[FLAC_Model_Predictions$UTI==0,
                                    c("FLAC_Prob")]
FLAC_Prob_High <- FLAC_Model_Predictions[FLAC_Model_Predictions$UTI==1,
                                    c("FLAC_Prob")]
FLAC_Model_Prob_ROC <- roc.curve(scores.class1 = FLAC_Prob_Low,
                               scores.class0 = FLAC_Prob_High,
                                curve = TRUE)

plot(FLAC_Model_Prob_ROC,
     xlab="1-Specificity", 
     ylab="Sensitivity",
     main="ROC Curve of the UTI Probabilities : Firth's Logistic Regression With Added Covariate", 
     color=TRUE, 
     lwd=8,
     legend=3)

```

###  1.5.4 Firth's Logistic Regression With Intercept Correction (FLIC)

```{r section_1.5.4, warning=FALSE, message=FALSE}

```

###  1.5.5 Bayesian Generalized Linear Model With Cauchy Priors (BGLMCP)

```{r section_1.5.5, warning=FALSE, message=FALSE}

```

###  1.5.6 Ridge Regression (RR)

```{r section_1.5.6, warning=FALSE, message=FALSE}

```


##  1.6 Probability Curve and Coefficient Estimation Evaluation Summary
|
| Model estimation comparison:
|
```{r section_1.6, warning=FALSE, message=FALSE}

```

# **2. References**
|
| **[Book]** [Applied Predictive Modeling](http://appliedpredictivemodeling.com/) by Max Kuhn and Kjell Johnson
| **[Book]** [An Introduction to Statistical Learning](https://www.statlearning.com/) by Gareth James, Daniela Witten, Trevor Hastie and Rob Tibshirani
| **[Book]** [Multivariate Data Visualization with R](http://lmdvr.r-forge.r-project.org/figures/figures.html) by Deepayan Sarkar
| **[Book]** [Machine Learning](https://bookdown.org/ssjackson300/Machine-Learning-Lecture-Notes/) by Samuel Jackson
| **[Book]** [Data Modeling Methods](https://bookdown.org/larget_jacob/data-modeling-methods/) by Jacob Larget
| **[Book]** [Introduction to R and Statistics](https://saestatsteaching.tech/) by University of Western Australia
| **[Book]** [Feature Engineering and Selection: A Practical Approach for Predictive Models](http://www.feat.engineering/index.html) by Max Kuhn and Kjell Johnson
| **[Book]** [Introduction to Research Methods](https://bookdown.org/ejvanholm/Textbook/) by Eric van Holm
| **[R Package]** [AppliedPredictiveModeling](https://cran.r-project.org/web//packages/AppliedPredictiveModeling/AppliedPredictiveModeling.pdf) by Max Kuhn
| **[R Package]** [caret](https://topepo.github.io/caret/index.html) by Max Kuhn
| **[R Package]** [rpart](https://mran.microsoft.com/web/packages/rpart/rpart.pdf) by Terry Therneau and Beth Atkinson
| **[R Package]** [lattice](https://cran.r-project.org/web/packages/lattice/lattice.pdf) by  Deepayan Sarkar
| **[R Package]** [dplyr](https://cran.r-project.org/web/packages/dplyr/index.html/) by Hadley Wickham
| **[R Package]** [moments](https://cran.r-project.org/web/packages/moments/index.html) by Lukasz Komsta and Frederick
| **[R Package]** [skimr](https://cran.r-project.org/web/packages/skimr/skimr.pdf) by  Elin Waring
| **[R Package]** [RANN](https://cran.r-project.org/web/packages/RANN/RANN.pdf) by  Sunil Arya, David Mount, Samuel Kemp and Gregory Jefferis
| **[R Package]** [corrplot](https://cran.r-project.org/web/packages/corrplot/corrplot.pdf) by Taiyun Wei
| **[R Package]** [tidyverse](https://cran.r-project.org/web/packages/tidyverse/tidyverse.pdf) by Hadley Wickham
| **[R Package]** [lares](https://cran.rstudio.com/web/packages/lares/lares.pdf) by Bernardo Lares
| **[R Package]** [DMwR](https://mran.microsoft.com/snapshot/2016-05-02/web/packages/DMwR/DMwR.pdf) by Luis Torgo
| **[R Package]** [gridExtra](https://cran.r-project.org/web/packages/gridExtra/gridExtra.pdf) by Baptiste Auguie and Anton Antonov
| **[R Package]** [rattle](https://cran.r-project.org/web/packages/rattle/rattle.pdf) by Graham Williams
| **[R Package]** [rpart.plot](https://cran.r-project.org/web/packages/rpart.plot/rpart.plot.pdf) by Stephen Milborrow
| **[R Package]** [RColorBrewer](https://cran.r-project.org/web//packages/RColorBrewer/RColorBrewer.pdf) by Erich Neuwirth
| **[R Package]** [stats](https://search.r-project.org/R/refmans/stats/html/00Index.html) by R Core Team
| **[R Package]** [pls](https://cran.r-project.org/web/packages/pls/pls.pdf) by Kristian Hovde Liland
| **[R Package]** [nnet](https://cran.r-project.org/web/packages/nnet/nnet.pdf) by Brian Ripley
| **[R Package]** [elasticnet](https://cran.r-project.org/web/packages/elasticnet/elasticnet.pdf) by Hui Zou
| **[R Package]** [earth](https://cran.r-project.org/web/packages/earth/earth.pdf) by Stephen Milborrow
| **[R Package]** [party](https://cran.r-project.org/web/packages/party/party.pdf) by Torsten Hothorn
| **[R Package]** [kernlab](https://cran.r-project.org/web/packages/kernlab/kernlab.pdf) by Alexandros Karatzoglou
| **[R Package]** [randomForest](https://cran.r-project.org/web/packages/randomForest/randomForest.pdf) by Andy Liaw
| **[R Package]** [pROC](https://cran.r-project.org/web/packages/pROC/pROC.pdf) by Xavier Robin
| **[R Package]** [mda](https://cran.r-project.org/web/packages/mda/mda.pdf) by Trevor Hastie
| **[R Package]** [klaR](https://cran.r-project.org/web/packages/klaR/klaR.pdf) by Christian Roever, Nils Raabe, Karsten Luebke, Uwe Ligges, Gero Szepannek, Marc Zentgraf and David Meyer
| **[R Package]** [pamr](https://cran.r-project.org/web/packages/pamr/pamr.pdf) by Trevor Hastie, Rob Tibshirani, Balasubramanian Narasimhan and Gil Chu
| **[R Package]** [OptimalCutpoints](https://cran.r-project.org/web/packages/OptimalCutpoints/OptimalCutpoints.pdf) by Monica Lopez-Raton and Maria Xose Rodriguez-Alvarez
| **[R Package]** [broom](https://cran.r-project.org/web/packages/broom/broom.pdf) by Simon Couch
| **[R Package]** [PRROC](https://cran.r-project.org/web/packages/PRROC/PRROC.pdf) by Jan Grau and Jens Keilwagen
| **[R Package]** [logistf](https://cran.r-project.org/web/packages/logistf/logistf.pdf) by Georg Heinze
| **[R Package]** [arm](https://cran.r-project.org/web/packages/arm/arm.pdf) by Yu-Sung Su
| **[R Package]** [glmnet](https://cran.r-project.org/web/packages/glmnet/glmnet.pdf) by Trevor Hastie
| **[Article]** [The caret Package](https://topepo.github.io/caret/index.html) by Max Kuhn
| **[Article]** [A Short Introduction to the caret Package](https://cran.r-project.org/web/packages/caret/vignettes/caret.html) by Max Kuhn
| **[Article]** [Caret Package – A Practical Guide to Machine Learning in R](https://www.machinelearningplus.com/machine-learning/caret-package/#:~:text=Caret%20is%20short%20for%20Classification%20And%20REgression%20Training.,track%20of%20which%20algorithm%20resides%20in%20which%20package.) by Selva Prabhakaran
| **[Article]** [Tuning Machine Learning Models Using the Caret R Package](https://machinelearningmastery.com/tuning-machine-learning-models-using-the-caret-r-package/) by Jason Brownlee
| **[Article]** [Lattice Graphs](http://www.sthda.com/english/wiki/lattice-graphs) by Alboukadel Kassambara

| **[Article]** [What is Complete or Quasi-Complete Separation in Logistic/Probit Regression and How Do We Deal with Them](https://stats.oarc.ucla.edu/other/mult-pkg/faq/general/faqwhat-is-complete-or-quasi-complete-separation-in-logisticprobit-regression-and-how-do-we-deal-with-them/) by UCLA Advanced Research Computing Group
| **[Article]** [What Is Complete Separation in Binary Logistic Regression?](https://blog.minitab.com/en/starting-out-with-statistical-software/what-is-complete-separation-in-binary-logistic-regression) by Eric Heckman 
| **[Article]** [xxx](xxx) by xxx 
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Article]** [xxx](xxx) by xxx
| **[Publication]** [Separation in Logistic Regression: Causes, Consequences, and Control](https://academic.oup.com/aje/article/187/4/864/4084405?login=false) by Mohammad Ali Mansournia, Angelika Geroldinger, Sander Greenland and Georg Heinze (American Journal of Epidemiology)
| **[Course]** [Applied Data Mining and Statistical Learning](https://online.stat.psu.edu/stat508/) by Penn State Eberly College of Science
|
|
|
|